Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-15 thesis/paper_draft.qmd
finished the contribution part in intro section, and write conclusion, limitation/future works

In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (the sum of all countries' sales does not equal the total sales). Such forecasts are called *incoherent*. First, they undermine downstream decisions that require internal consistency, and second, our forecast performance suffers. To overcome these issues, forecast reconciliation was introduced. Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting forecasts are *coherent* (i.e., respecting the aggregation constraints). It first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others to enhance point forecast performance. Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq provides good theoretical properties of minimising the total variance of the reconciled forecast errors, fast computational speed, and significant empirical improvements in many applications. It has become a standard method for forecast reconciliation, implemented in popular R and Python softwares/packages [@O-Hara-Wild2024-we; @Nixtla].

@@

The theorectical advantage of MinT might not be achieved in finite sample settings if the covariance estimate is not close enough to the true covariance matrix. Despite its importance, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who proposed a Double Shrinkage estimator for MinT. The idea is to introduce a second shrinkage target, which incorporates conditional dependence information suggested by the hierarchical structure. This paper is still in its early stage, leaving many potential gaps still not yet to explore.

@@

The purpose of this paper is to assess the forecasting performance of MinT when different covariance estimators are used. First, we show

The remainder of the paper is organised as follows. @sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Min Trace approach, introducing notations, terminologies, and motivations for alternative estimators. @sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses. @sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT. @sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.

@@

The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].

@@

For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set. In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts. We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).

@@

>>>>

In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (the sum of all countries' sales does not equal the total sales). Such forecasts are called *incoherent*. First, they undermine downstream decisions that require internal consistency, and second, our forecast performance suffers. To overcome these issues, forecast reconciliation was introduced. Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting forecasts are *coherent* (i.e., respecting the aggregation constraints). It first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others to enhance point forecast performance. @Athanasopoulos2024-as provided a comprehensive review of the literature on forecast reconciliation. Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq provides good theoretical properties of minimising the total variance of the reconciled forecast errors, fast computational speed, and significant empirical improvements in many applications. It has become a standard method for forecast reconciliation, implemented in popular R and Python softwares/packages [@O-Hara-Wild2024-we; @Nixtla].

@@

The theorectical advantage of MinT might not be achieved in finite sample settings if the covariance estimate is not close enough to the true covariance matrix. Despite its importance, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who proposed a Double Shrinkage estimator for MinT. The idea is to introduce a second shrinkage target, which incorporates conditional dependence information suggested by the hierarchical structure. This paper is still in its early stage, and does not fully address the shortcomings mentioned above.

@@

The purpose of this paper is to explore the shortcomings of the shrinkage estimator and introduce alternative estimators that address these issues one by one. We evaluate the performance of MinT with these different covariance estimators in both point and probabilistic forecast reconciliation settings. In our empirical results from the complex, large-hierarchy dataset, we observe the following: (i) MinT with NOVELIST estimator with brings improvements over the shrinkage estimator under the probabilistic setting, as its predictive distributions have higher coverage rates; (ii) Covariance estimators that utilise the latent factors structure consistently produce better point and probabilistic forecasts than those that do not, when the data is evidently driven by such structure; (iii) Scaling the 1-step-ahead covariance matrix to approximate the h-step-ahead one gives significantly better performance than estimators that do not use this assumption, suggesting that the scaling assumption is not entirely incorrect. The estimators relaxing on the proportionality assumption that we propose can also be further extended to h-step-ahead probabilistic forecast reconciliation.

The remainder of the paper is organised as follows. @sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Minimum Trace approach, introducing notations, terminologies, and motivations for alternative estimators. @sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses. @sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT. @sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.

@@

<!-- The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov]. -->

@@

For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set. In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts. We then reconcile these forecasts under different methods, and evaluate their point and probabilistic forecast accuracy on the test set.

@@

# Conclusions and Future Work {#sec-conclusion}

<!-- 
- Aim to study the shortcomings of current MinT, a method attracting lots of attention in academia and practice.
- Propose a NOVELIST, PC-adjusted variants, and multi-step covariance variants.
- Empirical results on tourism data show that:
  - MinT with shrinkage is a solid default for point forecasts. NOVELIST performs more reliably for probabilistic forecasts.
  - PC adjustment with a single dominant factor consistently enhances performance in both point and probabilistic forecast and should be considered when a dominant latent factor is evident.
  - Proportionality assumption still gives reasonably good results.
  - More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
- Limitations & Future works:
  - More complex simulation studies needed to tease out the differences between shrinkage and NOVELIST. No simulating from factor models.
  - Not standardising the data for PC-variants methods, higher-level series can dominate the PCs.
  - Explore other methods utilising cross-series information, eigenvalue/eigenvector structures,..
  - Explore why MinT shrinkage fails for Winkler 95% score, especially at higher levels.
  - h-step-ahead probabilistic forecast evaluation.
-->

This paper aims to address the limitations of the current Minimum Trace reconciliation method, a method that has garnered significant attention in both academic and practical forecasting contexts. We propose the NOVELIST estimator to address the lack of flexibility in the uniform shrinkage target of the traditional shrinkage estimator. Additionally, we introduce PC-adjusted variants of shrinkage and NOVELIST to take into account dominant latent factors, as well as multi-step covariance variants that relax the proportionality assumption.

Empirical results on Australian domestic tourism data reveal several insights. For point forecasts, shrinkage (and its PC-adjusted version) remains a robust default choice, while NOVELIST demonstrates improved performance for probabilistic forecasts. Adjusting for a single dominant principal component consistently enhances performance across both point and probabilistic forecasts, suggesting its utility when such latent structures are present. More complex adjustments, such as incorporating multiple principal components or horizon-specific covariances, tend to introduce additional estimation noise without notable benefits.

However, several limitations and avenues for future research remain. More intricate simulation studies are necessary to further highlight the differences between shrinkage and NOVELIST, particularly in scenarios involving factor models. The current implementation does not standardise data for PC-variant methods, which may lead to higher-level series largely influencing the principal components. Future work could also explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures. Additionally, understanding why MinT with shrinkage underperforms in the Winkler 95% score, especially at higher aggregation levels, needs further investigation. Finally, evaluating h-step-ahead probabilistic forecasts remains an open area for further research.
